### 2021-03-19 manuscript.Rmd
done for the night; figures look fine; need to finish last paragraph about filter choice

---
title: "Project 3: Comparison of variant calling by `freebayes` and `mpileup` from high-throughput sequencing of Burbot (*Lota lota*)"
author: "Jason Moggridge 1159229"
date: "2021/04/19"
output: 
  pdf_document:
    highlight: kate
  html_notebook:
    highlight: kate
bibliography: references.bib
link-citations: true
---

```{r setup, echo=F}
knitr::opts_chunk$set(warning = F, message = F, cache = T)
```

------------------------------------------------------------------------

# Introduction

The analysis of genetic variants such as single-nucleotide polymorphisms (SNPs) and insertions/deletions (indels) by high-throughput sequencing is a central objective in studies of quantitative-trait loci, genome-wide associations, population structure, among others. These polymorphic sites are also of practical interest in clinical genomics, *eg*. to discover disease-causing mutations [@cornish2015], and in agriculture, where SNPs can be used as genetic markers to guide breeding of crops [@yao2020] and livestock [@ni2015]. As the cost of sequencing has decreased greatly and computational resources have improved, these techniques have become increasingly popular.

The identification of polymorphic sites from sequencing data, known as variant calling, is performed after alignment of reads to a reference genome. Because high-throughput sequencing and read-mapping are noisy, it is challenging but essential to differentiate true variants from artefacts, particularly in low-coverage data [@nielsen2011]. Many variant calling tools (`bcftools`, `GATK`, `freebayes`, *etc.*) have been created for this classification task, though these differ in their error correction processes and statistical models for inferring genotypes. Further filtering of variants is generally applied after calling, and how this is performed depends on the nature of the study.

In the current work I compared variant calling by two tools: `bcftools mpileup/call` (previously `samtools`) and `freebayes`. The results were compared in terms of the number of SNPs, the overlap between sets, the depth at each position, and their minor allele frequencies. Both `bcftools` and `freebayes` apply Bayesian models to estimate genotype posterior probabilities, though they differ in how this is implemented. `bcftools` uses a two step process where `mpileup` first creates the 'pileup' of all reads at variant sites, and then `call` is used to infer genotypes at these [@li2011; @danecek2021]. This implementation has the advantage of being very fast. `freebayes` a uses different approach to inferring haplotypes in that it does not use the alignment but the full reads [@garrison2012; @howtof]. Previous comparisons of variant callers have often found low concordance between tools and issues with sensitivity [@orawe2013; @cornish2015]. However, others have reported high concordance and accuracy in their evaluation [@hwang2015]; the best method is likely dependent on nature of the sequencing data (*eg.* whole-genome or exome, RAD-seq, RNA-seq, *etc*.), the depth of coverage, and the type of variants of interest (*eg.* SNPs or indels).

# Methods

*All files can be found on `cedar` in `/scratch/jmoggrid/Project3/`.*

### Dataset

I used data from project \#2 in this work, specifically the reads mapped to the burbot reference with the `bowtie2` aligner. First, I needed to convert my `.sam` files to `.bam` format and index them, as well as adding read groups using `picard` (necessary for `freebayes`):

<!-- fyi - bash code chunks have engine='bash' in their header -->

```{r sam-to-bam and readgroups, engine = 'bash', eval = FALSE, echo=TRUE}
#!/bin/bash
# to convert .sam to .bam, then sort & index
module load samtools
for file in bam/*.sam; do
  name=`echo $file | sed 's/\.sam//'`
  samtools view -b -S -o $name.bam $file
  samtools sort $name.bam -o $name.sorted.bam;
  samtools index $name.sorted.bam
done

# add readgroups and re-index; show output to verify
module load picard 
group=1
for file in bam/*.sorted.bam
do
  label=`echo $file | sed 's/.sorted.bam//'`
  java -jar $EBROOTPICARD/picard.jar AddOrReplaceReadGroups I=$file \
    O=$label.rg.sorted.bam RGID=$group RGLB=lib1 RGPL=illumina RGPU=unit1 RGSM=$label
  samtools index $label.rg.sorted.bam
  echo "group number: $group; file: $file; read group:"
  samtools view -H $label.rg.sorted.bam | grep "@RG"
  group=$((group+1))
done
```

### `bcftools mpileup` and `freebayes` variant calling

I used `bcftools mpileup/call` (referred to as `mpileup`) and `freebayes` to call variants from the ten burbot alignments. For `mpileup`, I added arguments to include annotations for read depth and allelic depth (-a DP,AD) and to specify Illumina as the sequencing platform (-P). For `call`, I specified the use of the newer multi-allelic caller (m), 'variants only' (v), to skip indels, and to provide the genotype quality scores (GQ). A summary of the raw SNP calls was obtained using `bcftools stats`.

```{r mpileup, engine = 'bash', eval = FALSE, echo=TRUE}
module load nixpkgs/16.09  gcc/7.3.0 bcftools/1.9
# mpileup and variant calling
bcftools mpileup -Ou -a DP,AD -P ILLUMINA \
-f ./burbot_ref/GCA_900302385.1_ASM90030238v1_genomic.fna ./bam/*.rg.sorted.bam | \
bcftools call -mv --format-fields GQ --skip-variants indels -o  mpileup.call.vcf
bcftools stats mpileup.call.vcf > mpileup.call.stats
```

I called SNPs using `freebayes` with default settings except for the flag to provide annotations for genotype quality. Output was limited to SNPs with `vcffilter`.

```{r freebayes, engine = 'bash', eval = FALSE, echo = TRUE}
module load freebayes/1.2.0
# create a list of 10 filenames to pass to as input to freebayes
> files.txt
for file in bam/*.rg.sorted.bam; do echo $file >> files.txt; done
# free-bayes calling for 10 files and restricting vcf output to SNPs w vcffilter
freebayes -f burbot_ref/GCA_900302385.1_ASM90030238v1_genomic.fna  \
  --genotype-qualities --bam-list files.txt | \
vcffilter -f "TYPE = snp" > freebayes.call.vcf
bcftools stats freebayes.call.vcf > freebayes.call.vcf
```

### Filtering variant calls

I used the `vcftools` library to filter variant calls. Biallelic SNPs (`--min-alleles 2 --max-alleles 2 --remove-indels`) were retained if they had a quality score of at least 20 (`--minQ 20`). Genotypes were retained at these sites if they had genotyping quality score of at least 10 (`--minGQ 10`). I found the vast majority of variants to be shared among all 10 individuals (data not shown), so I applied filters for the maximum and minimum minor allele frequencies, such that these would not be either zero or one (`–maf 0.01 –max-maf 0.99`). Any calls that did not receive a 'PASS' flag were discarded (`--remove-filtered-all`).

```{r filtering, engine='bash', eval = FALSE, echo = TRUE}
module --force purge; module load StdEnv/2020 vcftools/0.1.16

# filter mpileup calls 
vcftools --vcf mpileup.call.vcf --minQ 20 --minGQ 10 \
  --min-alleles 2 --max-alleles 2 --maf 0.01 --max-maf 0.99 \
  --remove-filtered-all --recode --recode-INFO-all \
  --out mpileup.filter

# same thing for freebayes
vcftools --vcf freebayes.call.vcf --minQ 20 --minGQ 10 \
  --min-alleles 2 --max-alleles 2 --maf 0.01 --max-maf 0.99 \
  --remove-filtered-all --recode --recode-INFO-all \
  --out freebayes.filter
```

There was no need to exclude any individuals from the analysis, as none had more that 22.5% missing data in the filtered `mpileup` calls, while the proportions of missing data in `freebayes` calls was roughly half as large (from `--missing-indv`).

```{r check_missing, engine='bash', eval = FALSE, echo = TRUE}
# check individuals' missing data
vcftools --vcf mpileup.filter.recode.vcf --missing-indv \
  --out mpileup.filter
vcftools --vcf freebayes.filter.recode.vcf --missing-indv \
  --out freebayes.filter
```

### Summarizing SNP calls

I counted the biallelic SNP calls from each caller with `bcftools stats`.

```{r count_snps, engine = 'bash', eval = FALSE, echo = TRUE}
module load StdEnv/2020 bcftools/1.10.2
bcftools stats mpileup.filter.recode.vcf | grep "number of SNPs:" | cut -f4
bcftools stats freebayes.filter.recode.vcf | grep "number of SNPs:" | cut -f4
```

The overlap between sets of SNPs was computed with `bcftools isec`.

```{r overlap_in_sets, engine = 'bash', eval = FALSE, echo = TRUE}
# prepare files for isec by zipping and re-indexing
bgzip mpileup.filter.recode.vcf
bcftools index mpileup.filter.recode.vcf.gz
bgzip freebayes.filter.recode.vcf
bcftools index freebayes.filter.recode.vcf.gz
# Compute intersection of 2 files with isec
bcftools isec -p isec mpileup.filter.recode.vcf.gz freebayes.filter.recode.vcf.gz
# Get overlap data from isec into tsv format
> Overlap.vals
for file in isec/*.vcf
do
  bcftools stats $file | grep "number of SNPs:" | cut -f4 >> Overlap.vals
done
echo $'unique to mpileup\nunique to freebayes\nshared\nshared' > Overlap.names
paste  Overlap.names Overlap.vals > Overlap.tsv
sed -i '1s/^/SNPs\tCount\n/' Overlap.tsv
```

Using `vcftools`, I computed the allele frequency at each loci across individuals, as well as the depth at each locus summed across individuals.

```{r vcf_freq, engine = 'bash', eval = FALSE, echo = TRUE}
# allele frequency
vcftools --gzvcf mpileup.filter.recode.vcf.gz --freq  --out mpileup.filter
vcftools --gzvcf freebayes.filter.recode.vcf.gz --freq --out freebayes.filter
# site depth sum
vcftools --gzvcf mpileup.filter.recode.vcf.gz --site-depth --out mpileup
vcftools --gzvcf freebayes.filter.recode.vcf.gz --site-depth --out freebayes
```

### Statistics

I created figures and summary tables from the data generated above using the `R` language. *This code is presented in the supplementary section at the end of the manuscript.*

```{r load data, message=FALSE, echo=FALSE}
#! {R code}
library(tidyverse, quietly = T)

# Overlap counts table; add totals for freebayes and mpileup 
Overlap <- read_delim('./data/Overlap.tsv', '\t') %>%  
  distinct() %>% 
  bind_rows(
    tibble(SNPs = c('mpileup', 'freebayes'), 
           Count = c(.[1,2][[1]] + .[3,2][[1]], 
                     .[2,2][[1]] + .[3,2][[1]]))) %>%  
  arrange(desc(Count))
```

```{r load_data2, message=FALSE, echo = F}
# Allele frequencies
cnames <- c('chrom','pos', 'n_alleles','n_chr','a1','a2')
AlleleFreq <-  
  bind_rows(
    read_tsv('./data/freebayes.filter.frq', col_names = cnames, skip = 1L) %>% 
      mutate(caller = 'freebayes'),
    read_tsv('./data/mpileup.filter.frq', col_names = cnames, skip = 1L) %>% 
      mutate(caller = 'mpileup')
  ) %>% 
  mutate(allele = str_extract(a1, '[A-Z]+'), 
         frequency = str_remove_all(a1, '[A-Z]+:'),
         allele2 = str_extract(a2, '[A-Z]+'), 
         frequency2 = str_remove_all(a2, '[A-Z]+:')
  ) %>% 
  mutate(across(contains('freq'), as.numeric)) %>%
  mutate(
    maf = ifelse(frequency > 0.5, 1 - frequency, frequency),
    minor_allele = ifelse(maf == frequency, allele, allele2)
  )

# Depth
cnames <- c('individual', 'n sites', 'mean_depth')
IndvDepth <- bind_rows(
    read_tsv('./data/freebayes.idepth', col_names = cnames, skip = 1L) %>%
      mutate(caller = 'freebayes'),
    read_tsv('./data/mpileup.idepth', col_names = cnames, skip = 1L) %>%
      mutate(caller = 'mpileup')
  ) %>% 
  mutate(individual = str_remove_all(individual, 'bam.|.bowtie.burbot'),
         caller = as_factor(caller)) %>% 
  arrange(individual)

# Site Depth
cnames <- c("chr", "pos", "allele_depth", "var_depth")
SiteDepth <- bind_rows(
  read_tsv('./data/freebayes.ldepth', col_names = cnames, skip = 1L) %>%
    mutate(caller = 'freebayes'),
  read_tsv('./data/mpileup.ldepth', col_names = cnames, skip = 1L) %>%
    mutate(caller = 'mpileup')
) %>% mutate(caller = as_factor(caller))

vcf_df <- SiteDepth %>%
  rename(chrom = chr) %>% 
  left_join(AlleleFreq, by = c('chrom', 'pos', 'caller'))

cnames <- c('id', 'quality', 'n_snps', 'ts', 'tv', 'indels')
qual_df <- 
  bind_rows(
    read_tsv("./data/mpileup.quality.stats", skip = 1L, col_names = cnames) %>%  mutate(caller = "mpileup"),
    read_tsv("./data/freebayes.quality.stats", skip = 1L, col_names = cnames) %>%  mutate(caller = "freebayes")
  ) %>%  mutate(caller = as_factor(caller))
```

```{r figs, echo=F}
library(ggbeeswarm)
library(patchwork)
library(rcartocolor)
library(gridExtra)
library(grid)

# Allele frequency histogram (alleles stacked)
fig_allelefreq <- 
  ggplot(AlleleFreq) +
  geom_histogram(aes(x = maf), bins = 20, alpha = 0.8) +
  facet_wrap(~caller, scales = 'free_y', nrow = 2) +
  labs(x = 'Minor Allele Frequency', y='Number of Loci') +
  theme_bw()

# Allele depth over maf, coloured by n chromosomes
fig_alleledepth_over_maf <- 
  ggplot(vcf_df, aes(x=maf, y=allele_depth, color = n_chr)) +
  geom_point(size = 0.5, alpha = 0.5) +
  scale_y_log10() +
  facet_wrap(caller~., scales = 'free_y', nrow = 2) +
  scale_color_viridis_c() +
  labs(x = "Minor allele frequency",
       y = "Depth", 
       color="Number of\nchromosomes") +
  theme_bw()

# The distribution of depth per-site summed across individuals
fig_sitedepth <- 
  ggplot(SiteDepth, aes(x = allele_depth, fill = caller, color = caller)) +
  geom_density(alpha = 0.12) +
  geom_vline(aes(xintercept = mean(allele_depth), color = caller),
             lty = 2) +
  scale_x_log10() +
  labs(x = 'Depth', y = 'Density') +
  guides(colour = guide_legend(override.aes = list(size=1, pch=15))) +
  theme_bw()
  # theme(legend.position = c(.8,.7))

# The distribution of mean depth per site within individuals
set.seed(45)
fig_indvdepth <-
  ggplot(IndvDepth, aes(y = mean_depth, x = caller)) +
  geom_violin(alpha = 0.1, draw_quantiles = c(.25, .5, .75), trim = F) +
  geom_quasirandom(aes(color = individual), size = 3, width = 0.3) +
  scale_color_carto_d(palette = 2) +
  guides(colour = guide_legend(override.aes = list(size = 4, pch = 16))) +
  facet_wrap( ~ caller, scales = 'free_x') +
  labs(y = 'Mean Depth', x = '', color = 'Subject') +
  theme_bw() +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    legend.position = 'right'
  )
fig_quality_dist <- qual_df %>% 
  mutate(QUAL = round(quality/25)*25) %>% 
  group_by(QUAL, caller) %>% 
  summarise(n_snps = sum(n_snps),
            ts  = sum(ts),
            tv = sum(tv)) %>% 
  pivot_longer(cols = c(ts,tv)) %>% 
  ggplot(aes(x = QUAL, y = value, fill = name)) +
  geom_col(width = 19) +
  facet_wrap(~caller) +
  scale_fill_carto_d(palette = 2) +
  labs(x="Quality Score", y = "Number of Loci", fill = NULL) +
  scale_y_log10() +
  theme_bw()
```

------------------------------------------------------------------------

# Results

From the ten burbots in our sample, mpileup called a total of 132,046 SNPs, 173 of which were multiallelic, and with a transition:transversion ratio (ts:tv) of 0.81; freebayes called a total of 106,973 SNPs, with 1761 being multiallelic, and ts:tv of 0.44 (table 1). After applying filters for these were reduced to 11,115 SNPs with a 0.87 transition:transversion ratio for mpileup calls and 6,722 SNPs with 1.09 transition:transversion ratio for freebayes calls. The sets of biallelic SNPs identified by each caller were compared (table 2): mpileup called `r round(Overlap[1,2][[1]]* 100 / Overlap[2,2][[1]], 2) - 100` % more biallelic SNPs than freebayes. Of these, a large proportion (`r scales::comma(Overlap[3,2][[1]])`) were shared between both sets, representing `r round(Overlap[3,2][[1]]*100 /Overlap[1,2][[1]], 2)` % of mpileup calls and `r round(Overlap[3,2][[1]]*100 /Overlap[2,2][[1]], 2)` % of freebayes calls. Additionally, `r Overlap[4,2][[1]]` SNPs were unique to mpileup and `r Overlap[5,2][[1]]` unique to freebayes.

```{r table1_filtering, echo=FALSE}
tribble(
  ~ Caller, ~ `raw SNPs`, ~ `raw ts/tv`, ~ `post-filter SNPs`, ~ `post-filter ts/tv`,
  "mpileup", '132,046', 0.81, '11,115', 0.87,
  "freebayes", '106,973', 0.44, '6,722', 1.09
) %>% 
  pander::pander(caption = "Raw and filtered SNP counts and transition/transversion ratios for calls by mpileup and freebayes.")
```

The distribution minor allele frequencies of SNPs from mpileup shows a U-shape with peaks at 0.05 and 0.5, whereas those from freebayes peak at 0.05 and are otherwise evenly distributed (fig.1A). The range of possible values is constrained due to the small sample size (some values can only be obtained with a specific amount of missing data; fig. 1B). The sum depth per locus across all individuals is generally in the range of 75x-200x, with both callers showing similar distributions (fig. 2A). The mean depth for mpileup was 174 reads and for freebayes 162 reads. Both callers have some extreme outliers in terms of site depth: the maximum depth for freebayes was 16,162 reads, while for mpileup it was only 2522 reads. The minimum site depth was 5 and 3 for freebayes and mpileup respectively and mpileup has a greater proportion of low-depth sites in general. Freebayes also had a greater mean depth per individual (fig. 2B). Both callers show similar distributions of quality scores, though mpileup calls a greater proportion of transversions at low-quality loci than freebayes (fig. 2C).

```{r table2_overlap, echo=F}
Overlap %>% 
  mutate(Count = scales::comma(Count)) %>% 
  pander::pander(caption = "Summary of total, shared, and unique SNPs called by mpileup and freebayes (after filtering).")
```

```{r fig1, echo=F, fig.align = 'center', fig.width=8, fig.height=4, fig.cap="(A) Distribution of minor allele frequency; SNPs with frequency = 0 or 1 were omitted. (B) Distribution of depth per locus summed across all individuals (nb: the x-axis is log-scaled)."}
(fig_allelefreq + fig_alleledepth_over_maf) &
  plot_layout(guides = 'keep') &
  plot_annotation(tag_levels = 'A')
```

```{r fig2, echo=F, fig.align = 'center', fig.width=6, fig.height=8, fig.cap="(A) Scatterplot of loci (points) by allele depth and minor allele frequency, colored by the number of chromosomes counted. Due to the sample size (n=10), minor allele frequencies are constrained by the number of chromosomes at the SNP site; there appears to be no relationship between depth (ie. read coverage) and minor allele frequencies. \n(B) Mean depth per locus for each individual. Violin plot lines indicate quartiles. \n(C) The quality score distributions for each set of SNP calls, separated into transitons (ts) and transversions (tv)"}
 fig_sitedepth / fig_indvdepth / fig_quality_dist & 
  plot_annotation(tag_levels = 'A')
```

------------------------------------------------------------------------

# Discussion

In this work I compared the sets of biallelic SNPs called by two tools, bcftools mpileup/call and freebayes, from alignments of burbot reads from 10 individuals. The raw SNP calls were filtered similarly for both sets based on quality scores, genotyping confidence scores, and minor allele frequency; most of the data (\>90%) was discarded during filtering (table 1), with roughly one-third of calls having allele frequency of 0 or 1 (*ie.* all individuals were similar but differed from the reference). Unexpectedly, I found that mpileup called many more SNPs that freebayes, and retained nearly twice as many after filtering. This is in contrast to findings from @cornish2015 who found that freebayes called more SNPs in human exome data, though both tools had similar-sized sets after filtering. The discrepancy between these findings does not seem to be explained by the slight differences in filtering (they filtered for read depth; I filtered for genotype quality), but could be related to differences between datasets [@hwang2015].

Interestingly, the sets of SNPs called by mpileup and freebayes showed a high degree of overlap, with over 90% of freebayes calls being shared with the larger mpileup set (table 2). This result is somewhat in agreement with @hwang2015, who found \~90% concordance between (Samtools) mpileup, freebayes, and GATK on the gold-standard Genome in a Bottle dataset. My results suggest then, that either mpileup has low specificity (calls variants that don't exist), or that freebayes has low sensitivity (misses true variants). This does not seem to be the case generally though, as @cornish2015 suggest the opposite and found that a variety of tools had similarly low sensitivity on the GiaB gold-standard data.

My analysis also revealed differences in the distribution of minor allele frequencies (MAF) and read depth at SNP loci (fig. 1). Specifically, mpileup shows a clear preference for SNPs with MAF of 0.5 and to a lesser degree 0.25 (relative to similar values). Both callers have a peak at low MAF (0.05) but otherwise mpileup has an even distribution from 0.10 to 0.5, when accounting for the constraint on these values by the small sample size (fig. 1B). The callers show differences in the read depth across loci (fig. 2A) and individuals (fig. 2B), as high-depth sites (\>250x/ind.) are hard-filtered by `mpileup`, and mpileup identifies many more low-depth sites (\<70x) that are ignored by freebayes. As such, the concordance between sets would likely increase if stricter filters were used to mitigate these biases.

Certain characteristics of the burbot data present challenges for variant calling. The highly-fragmented assembly means we are likely to miss many variants since they are not represented in the alignment. The relatively low coverage of the data means that a number of true variants are likely discarded because of large uncertainty in their classification. Further, burbot is not a model organism and does not have the same resources to develop useful prior probabilities for variant calling, (*eg*. SNP databases for human variant calling; [@nielsen2011]), or benchmark data to validate pipelines (such as GiaB; [@zook2019a]).

In the present work, the small number of subjects means that the identification of uncommon and rare variants is highly unlikely. If we were more interested in finding such loci, we would want to apply less stringent filtering and accept the cost of increased false discoveries. However,

------------------------------------------------------------------------

\newpage

# Supplemental

I also examined the quality score distribution over sites and the mean per site depth by individual. These were computed with `bcftools stats` and `vcftools` respectively, as follows (bash code):

```{r stats_quality, engine = 'bash', eval=FALSE}
## quality scores from bf
grep -E "Quality" mpileup.filter.recode.stats | cut -f1 --complement > mpileup.quality.stats
grep ^QUAL mpileup.filter.recode.stats | cut -f1 --complement >> mpileup.quality.stats 
grep -E "Quality" freebayes.filter.recode.stats | cut -f1 --complement > freebayes.quality.stats
grep ^QUAL freebayes.filter.recode.stats | cut -f1 --complement >> freebayes.quality.stats 
# Mean depth for each individual across all loci
vcftools --gzvcf mpileup.filter.recode.vcf.gz --depth --out mpileup
vcftools --gzvcf freebayes.filter.recode.vcf.gz --depth --out freebayes
```

### R code for parsing results data

```{r supplementary1, eval = F, echo=T, include=T}
#! {R code}
library(tidyverse, quietly = T)
library(ggbeeswarm)
library(patchwork)
library(rcartocolor)
library(gridExtra)
library(grid)

# Overlap counts table; add totals for freebayes and mpileup 
Overlap <- read_delim('./data/Overlap.tsv', '\t') %>%  
  distinct() %>% 
  bind_rows(
    tibble(SNPs = c('mpileup', 'freebayes'), 
           Count = c(.[1,2][[1]] + .[3,2][[1]], 
                     .[2,2][[1]] + .[3,2][[1]]))) %>%  
  arrange(desc(Count))

# Allele frequencies
cnames <- c('chrom','pos', 'n_alleles','n_chr','a1','a2')
AlleleFreq <-  
  bind_rows(
    read_tsv('./data/freebayes.filter.frq', col_names = cnames, skip = 1L) %>% 
      mutate(caller = 'freebayes'),
    read_tsv('./data/mpileup.filter.frq', col_names = cnames, skip = 1L) %>% 
      mutate(caller = 'mpileup')
  ) %>% 
  mutate(allele = str_extract(a1, '[A-Z]+'), 
         frequency = str_remove_all(a1, '[A-Z]+:'),
         allele2 = str_extract(a2, '[A-Z]+'), 
         frequency2 = str_remove_all(a2, '[A-Z]+:')
  ) %>% 
  mutate(across(contains('freq'), as.numeric)) %>%
  mutate(
    maf = ifelse(frequency > 0.5, 1 - frequency, frequency),
    minor_allele = ifelse(maf == frequency, allele, allele2)
  )

# Depth
cnames <- c('individual', 'n sites', 'mean_depth')
IndvDepth <- bind_rows(
    read_tsv('./data/mpileup.idepth', col_names = cnames, skip = 1L) %>%
      mutate(caller = 'mpileup'),
    read_tsv('./data/freebayes.idepth', col_names = cnames, skip = 1L) %>%
      mutate(caller = 'freebayes')
  ) %>% 
  mutate(individual = str_remove_all(individual, 'bam.|.bowtie.burbot'),
         caller = as_factor(caller)) %>% 
  arrange(individual)

# Site Depth
cnames <- c("chr", "pos", "allele_depth", "var_depth")
SiteDepth <- bind_rows(
  read_tsv('./data/mpileup.ldepth', col_names = cnames, skip = 1L) %>%
    mutate(caller = 'mpileup'),
  read_tsv('./data/freebayes.ldepth', col_names = cnames, skip = 1L) %>%
    mutate(caller = 'freebayes')
) %>% mutate(caller = as_factor(caller))

vcf_df <- SiteDepth %>%
  rename(chrom = chr) %>% 
  left_join(AlleleFreq, by = c('chrom', 'pos', 'caller'))

cnames <- c('id', 'quality', 'n_snps', 'ts', 'tv', 'indels')
qual_df <- 
  bind_rows(
    read_tsv("./data/mpileup.quality.stats", skip = 1L, col_names = cnames) %>%  mutate(caller = "mpileup"),
    read_tsv("./data/freebayes.quality.stats", skip = 1L, col_names = cnames) %>%  mutate(caller = "freebayes")
  ) %>%  mutate(caller = as_factor(caller))
```

### R code for plotting results

```{r supplementary2, eval = F, echo=T, include=T}
## Figure 1L
# Allele frequency histogram (alleles stacked)
fig_allelefreq <- 
  ggplot(AlleleFreq) +
  geom_histogram(aes(x = maf), bins = 20, alpha = 0.8) +
  facet_wrap(caller~., scales = 'free_y') +
  labs(x = 'minor allele frequency', y='loci') +
  theme_bw()

# The distribution of depth per-site summed across individuals
fig_sitedepth <- 
  ggplot(SiteDepth, aes(x = allele_depth, fill = caller, color = caller)) +
  geom_density(alpha = 0.12) +
  geom_vline(aes(xintercept = mean(allele_depth), color = caller),
             lty = 2) +
  scale_x_log10() +
  labs(x = 'depth per locus', y = 'density') +
  guides(colour = guide_legend(override.aes = list(size=0.5, pch=15))) +
  theme_bw()

## Figure 2:

# Allele depth over maf, coloured by n chromosomes
fig_alleledepth_over_maf <- 
  ggplot(vcf_df, aes(x=maf, y=allele_depth, color = n_chr)) +
  geom_point(size = 0.5, alpha = 0.7) +
  scale_y_log10() +
  facet_wrap(~caller) +
  scale_color_viridis_c() +
  labs(x = "Minor allele frequency",
       y = "Allele depth", 
       color="Number of\nchromosomes") +
  theme_bw()

# The distribution of mean depth per site within individuals
set.seed(45)
fig_indvdepth <-
  ggplot(IndvDepth, aes(y = mean_depth, x = caller)) +
  geom_violin(alpha = 0.1, draw_quantiles = c(.25, .5, .75), trim = F) +
  geom_quasirandom(aes(color = individual), size = 3, width = 0.3) +
  scale_color_carto_d(palette = 2) +
  guides(colour = guide_legend(override.aes = list(size = 3, pch = 16))) +
  facet_wrap( ~ caller, scales = 'free_x') +
  labs(y = 'mean depth per locus', x = '', color = 'Individual') +
  theme_bw() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        legend.position = 'right')

# Quality score distribution; fill bars with type: ts or tv 
fig_quality_dist <- qual_df %>% 
  mutate(QUAL = round(quality/25)*25) %>% 
  group_by(QUAL, caller) %>% 
  summarise(n_snps = sum(n_snps),
            ts  = sum(ts),
            tv = sum(tv)) %>% 
  pivot_longer(cols = c(ts,tv)) %>% 
  ggplot(aes(x = QUAL, y = value, fill = name)) +
  geom_col(width = 19) +
  facet_wrap(~caller) +
  scale_fill_carto_d(palette = 1, direction = -1) +
  labs(x="quality score", y = "loci", fill = NULL) +
  scale_y_log10() +
  theme_bw()
```

### R code to output tables and figures

```{r supplementary3, eval = F, echo=T, include=T}
## table 1
tribble(
  ~ Caller, ~ `raw SNPs`, ~ `raw ts/tv`, ~ `post-filter SNPs`, ~ `post-filter ts/tv`,
  "mpileup", 132046, 0.81, 11115, 0.87,
  "freebayes", 106973, 0.44, 6722, 1.09 
  ) %>% 
  pander::pander(caption = "Raw and filtered SNP counts and
                 transition/transversion ratios for calls by mpileup 
                 and freebayes.")
### fig 1
(fig_allelefreq + fig_sitedepth) &
  plot_layout(guides = 'keep') &
  plot_annotation(tag_levels = 'A')
## table 2
Overlap %>% 
  pander::pander(caption = "Summary of total, shared, and unique SNPs
                 called by mpileup and freebayes (after filtering).")

### fig 2
(fig_alleledepth_over_maf / fig_indvdepth / fig_quality_dist) & 
  plot_annotation(tag_levels = 'A')
```

\newpage

# References